---
title: "PSTAT194CS_SNA_Final_Project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,   # hides code
  message = FALSE,
  warning = FALSE
)
```

# Dataset

# Loading and Preparing Data
```{r}
url = "https://snap.stanford.edu/data/web-Google.txt.gz"
download.file(url, destfile = "web-Google.txt.gz", mode = "wb")
```
```{r}
library(igraph)

edges = read.table(
  "web-Google.txt.gz",
  comment.char = "#",
)

graph = graph_from_data_frame(edges, directed=TRUE)
```

# Understanding our Network
```{r}
cat("Number of Nodes:",vcount(graph),"\n")
cat("Number of Edges:",ecount(graph))
```

```{r}
is_directed(graph)
```



```{r}
is_weighted(graph)
```


```{r}
summary(graph)
```






# Creating Subgraphs of 200-nodes



## Code Implemenation
```{r}
set.seed(123)

sub.network = lapply(1:4, function(x){
  induced_subgraph(graph, sample(V(graph),200))
})
```

## Checking Subgraphs
```{r}
sub.network[[1]]
sub.network[[2]]
sub.network[[3]]
sub.network[[4]]
```










## Random Walk New Implementation
```{r}
set.seed(123)

subgraphs <- list()

for(i in 1:4){
  start_node <- sample(V(graph), 1)
  walk <- random_walk(graph,
                      start = start_node,
                      steps = 500,
                      mode = "all")
  nodes_sampled <- unique(walk)
  nodes_sampled <- nodes_sampled[1:200]
  subgraphs[[i]] <- induced_subgraph(graph, nodes_sampled)
}
```

## Checking our Subgraphs
```{r}
subgraphs[[1]]
subgraphs[[2]]
subgraphs[[3]]
subgraphs[[4]]
```


## Checking on Sampling Bias
```{r}
deg.full = degree(graph)


bias.results = data.frame(
  Subgraph = character(),
  Nodes = numeric(),
  Edges = numeric(),
  Mean_Degree = numeric(),
  Median_Degree = numeric(),
  KS_p_value = numeric(),
  stringsAsFactors = FALSE
)

for (i in 1:4){
  deg.sample = degree(subgraphs[[i]])
  ks = ks.test(deg.sample, deg.full)
  bias.results = rbind(
    bias.results,
    data.frame(
      Subgraph = paste("Subgraph", i),
      Nodes = vcount(subgraphs[[i]]),
      Edges = ecount(subgraphs[[i]]),
      Mean_Degree = mean(deg.sample),
      Median_Degree = median(deg.sample),
      KS_p_value = ks$p.value
      )
    )
}
bias.results
```













# Visualization 

## Subgraph 1
```{r}
par(mfrow = c(1,3))

plot(
  subgraphs[[1]],
  layout = layout_with_fr(subgraphs[[1]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 1 - Fruchterman-Reingold"
)

plot(
  subgraphs[[1]],
  layout = layout_with_kk(subgraphs[[1]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 1 - Kamada-Kawai"
)

plot(
  subgraphs[[1]],
  layout = layout_with_mds(subgraphs[[1]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 1 - MDS"
)
```

## Subgraph 2
```{r}
par(mfrow = c(1,3))

plot(
  subgraphs[[2]],
  layout = layout_with_fr(subgraphs[[2]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 2- Fruchterman-Reingold"
)

plot(
  subgraphs[[2]],
  layout = layout_with_kk(subgraphs[[2]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 2 - Kamada-Kawai"
)

plot(
  subgraphs[[2]],
  layout = layout_with_mds(subgraphs[[2]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 2 - MDS"
)
```



## Subgraph 3
```{r}
par(mfrow = c(1,3))

plot(
  subgraphs[[3]],
  layout = layout_with_fr(subgraphs[[3]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 3 - Fruchterman-Reingold"
)

plot(
  subgraphs[[3]],
  layout = layout_with_kk(subgraphs[[3]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 3 - Kamada-Kawai"
)

plot(
  subgraphs[[3]],
  layout = layout_with_mds(subgraphs[[3]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 3 - MDS"
)
```




## Subgraph 4
```{r}
par(mfrow = c(1,3))

plot(
  subgraphs[[4]],
  layout = layout_with_fr(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - Fruchterman-Reingold"
)

plot(
  subgraphs[[4]],
  layout = layout_with_kk(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - Kamada-Kawai"
)

plot(
  subgraphs[[4]],
  layout = layout_with_mds(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - MDS"
)
```













# Metrics Evaluated for ALL 4 SUBGRAPHS (probably don't do this)

```{r, eval = TRUE}
for (i in 1:4) {
  cat("Metrics for Subgraph", i, ":\n", "NETWORK LEVEL METRICS", "\n")
  
  cat("Density:", "\n") 
  print(edge_density(subgraphs[[i]]))

  # Reciprocity
  cat("Reciprocity:", "\n")
  print(reciprocity(subgraphs[[i]]))

  # Dyad Census 
  cat("Pairwise Relationships:", "\n")
  print(dyad_census(subgraphs[[i]]))

  # Transitivity 
  cat("Transitivity:", "\n")
  print(transitivity(subgraphs[[i]]))
  
  # Average distance between nodes
  cat("Mean Distance Between Nodes:", "\n")
  print(mean_distance(subgraphs[[i]]))

  # Diameter
  cat("Diameter and the Associated Nodes:", "\n")
  print(diameter(subgraphs[[i]], weights = NA))
  print(farthest_vertices(subgraphs[[i]]))
  
  # NODE LEVEL METRICS
  cat("NODE LEVEL METRICS", "\n", "Degree:", "\n")
  print(degree(subgraphs[[i]], mode = "in"))
  
  # Centrality
  cat("Betweenness Centrality:", "\n")
  print(betweenness(subgraphs[[i]]))
  cat(mean(is.nan(betweenness(subgraphs[[i]]))), "of the values are NaN", "\n")
  
  cat("Closeness Centrality:", "\n")
  clr <- centr_clo(subgraphs[[i]], mode="all", normalized=T)
  print(centr_clo(subgraphs[[i]], mode="all", normalized=T))
  cat(mean(is.nan(closeness(subgraphs[[i]]))), "of the values are NaN", "\n")
  sort(clr$res, decreasing = T, index.return = T) # REMOVE ME Can remove if we don't want to plot
  
  cat("Local Transitivity:", "\n")
  print(transitivity(subgraphs[[i]], type="local"))

  # Edge Level
  edge_betweenness(subgraphs[[i]])
}
```


# 5) Community Detection

Because the graph is directed, we need to convert it into an undirected graph for community detection. 
```{r}
# convert all into undirected graphs
undirected_subgraphs <- vector("list", 4)
for (i in 1:4){
  undirected_subgraphs[[i]] <- as.undirected(subgraphs[[i]])
}
# list to store all results
all_results <- list()

for (i in 1:4) {
  g <- undirected_subgraphs[[i]]
  
  # run algs
  fg   <- cluster_fast_greedy(g)
  louv <- cluster_louvain(g)
  walk <- cluster_walktrap(g)
  
  # store into list
  all_results[[i]] <- data.frame(
    Subgraph      = i,
    Algorithm     = c("Fast Greedy", "Louvain", "Walktrap"),
    N_communities = c(length(fg), length(louv), length(walk)),
    Modularity_Q  = round(c(modularity(fg), modularity(louv), modularity(walk)), 4),
    Max_Comm_Size = c(max(sizes(fg)), max(sizes(louv)), max(sizes(walk)))
  )
}

# combine list into one table
final_comparison <- do.call(rbind, all_results)
final_comparison
```



# 6) Adjacency Matrix
```{r}
for(i in 1:4){
  adj_mat <- as_adjacency_matrix(subgraphs[[i]], sparse = FALSE)
  
  image(adj_mat,
        main = paste("Subgraph", i, "- Original Node Order"),
        col = c("white", "steelblue"),
        axes = FALSE)
}
# convert all into undirected graphs
undirected_subgraphs <- vector("list", 4)
for (i in 1:4){
  undirected_subgraphs[[i]] <- as.undirected(subgraphs[[i]])
}
# list to store all results
all_results <- list()

for (i in 1:4) {
  g <- undirected_subgraphs[[i]]
  
  # run algs
  fg   <- cluster_fast_greedy(g)
  louv <- cluster_louvain(g)
  walk <- cluster_walktrap(g)
  
  # store into list
  all_results[[i]] <- data.frame(
    Subgraph      = i,
    Algorithm     = c("Fast Greedy", "Louvain", "Walktrap"),
    N_communities = c(length(fg), length(louv), length(walk)),
    Modularity_Q  = round(c(modularity(fg), modularity(louv), modularity(walk)), 4),
    Max_Comm_Size = c(max(sizes(fg)), max(sizes(louv)), max(sizes(walk)))
  )
}

# combine list into one table
final_comparison <- do.call(rbind, all_results)
final_comparison
```



# 6) Adjacency Matrix
```{r}
for(i in 1:4){
  adj_mat <- as_adjacency_matrix(subgraphs[[i]], sparse = FALSE)
  
  image(adj_mat,
        main = paste("Subgraph", i, "- Original Node Order"),
        col = c("white", "steelblue"),
        axes = FALSE)
}
```



```{r}
set.seed(42)
for(i in 1:4){
  adj_mat <- as_adjacency_matrix(subgraphs[[i]], sparse = FALSE)
  
  kc_louv <- cluster_louvain(undirected_subgraphs[[i]])
  node_order <- order(membership(kc_louv))
  adj_mat_reordered <- adj_mat[node_order, node_order]
  
  image(adj_mat_reordered,
        main = paste("Subgraph", i, "- Reordered by Community"),
        col = c("white", "steelblue"),
        axes = FALSE)
}
```












>>>>>>> b78d14e34920b85431e3f00c1c3d5c71a504e95a





























\newpage

## Appendix

```{r appendix-code, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
