---
title: 'A Random Walk Through the Web: Network Analysis of the Google Web Graph'
author: "Richard Zhang, Toby Liu, Frank Wang, Max Peng"
subtitle: PSTAT 194CS
output:
  ioslides_presentation:
    widescreen: true
    smaller: true
  beamer_presentation: default
---

```{=html}
<style>
@import url('https://fonts.googleapis.com/css2?family=DM+Sans:wght@300;400;500;600&family=DM+Mono:wght@400;500&family=Playfair+Display:wght@600;700&display=swap');

/* ── Base ─────────────────────────────────────────────── */
slides > slide {
  background: #1c1f2b;
  color: #cbd3e8;
  font-family: 'DM Sans', sans-serif;
  font-size: 16px;
}

/* ── Title slide ──────────────────────────────────────── */
slides > slide.title-slide {
  background: linear-gradient(135deg, #1c1f2b 0%, #232840 50%, #1c1f2b 100%);
}

slides > slide.title-slide hgroup h1 {
  font-family: 'Playfair Display', serif;
  font-size: 2.6em;
  color: #eef1fa;
  margin-bottom: 0.2em;
}

slides > slide.title-slide hgroup h2 {
  font-family: 'DM Sans', sans-serif;
  font-weight: 300;
  color: #8fb3f0;
  font-size: 1.3em;
  letter-spacing: 0.04em;
}

/* ── Content slides ───────────────────────────────────── */
slides > slide:not(.title-slide) {
  background: #1c1f2b;
  border-top: 3px solid #4a7fd4;
  padding: 10px 50px 10px 50px;
}

slides > slide:not(.title-slide) > hgroup > h2 {
  font-family: 'Playfair Display', serif;
  color: #8fb3f0;
  font-size: 1.6em;
  border-bottom: 1px solid rgba(143, 179, 240, 0.3);
  padding-bottom: 5px;
  margin-bottom: 0.5em;
  margin-top: 0.3em;
}

/* ── Tables ───────────────────────────────────────────── */
table {
  width: 70%;
  border-collapse: collapse;
  font-family: 'DM Mono', monospace;
  font-size: 0.82em;
  margin: 0.4em auto;
}

th {
  background: #252a3d;
  color: #8fb3f0;
  font-weight: 500;
  padding: 7px 14px;
  text-align: left;
  letter-spacing: 0.04em;
  text-transform: uppercase;
  font-size: 0.78em;
}

td {
  padding: 6px 14px;
  border-bottom: 1px solid rgba(255,255,255,0.08);
  color: #6b7a9e;
}

tr:nth-child(even) td {
  background: rgba(74, 127, 212, 0.07);
}

/* ── Bullet points ────────────────────────────────────── */
ul, ol {
  margin-top: 0.3em;
}

li {
  margin-bottom: 0.4em;
  line-height: 1.6;
  color: #b8c4e0;
}

strong {
  color: #8fb3f0;
  font-weight: 600;
}

/* ── Inline code (backtick) ───────────────────────────── */
p code, li code {
  font-family: 'DM Mono', monospace;
  background: #252a3d;
  color: #7ecfa0;
  border: 1px solid rgba(126,168,232,0.2);
  border-radius: 4px;
  padding: 1px 6px;
  font-size: 0.87em;
}

/* ── Code blocks ──────────────────────────────────────── */
pre {
  font-family: 'DM Mono', monospace !important;
  background: #12151f !important;
  border: 1px solid rgba(143,179,240,0.18);
  border-radius: 6px;
  font-size: 0.79em;
  padding: 10px 14px;
  color: #b8d4a8 !important;
}

pre code {
  background: transparent !important;
  border: none;
  padding: 0;
  color: #b8d4a8 !important;
}

/* ── Slide number ─────────────────────────────────────── */
slides > slide:not(.title-slide)::after {
  color: #444d6a;
  font-size: 11px;
  font-family: 'DM Mono', monospace;
}

/* ── Logo banner ──────────────────────────────────────── */
.logo-banner {
  display: flex;
  justify-content: center;
  align-items: center;
  gap: 3em;
  margin: 1em auto 0.8em;
  padding: 0.9em 2em;
  background: #252a3d;
  border: 1px solid rgba(143,179,240,0.2);
  border-radius: 10px;
  width: fit-content;
}

.logo-banner img {
  height: 42px;
  object-fit: contain;
}

.logo-sep {
  width: 1px;
  height: 36px;
  background: rgba(143,179,240,0.3);
}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(igraph)
```





## Google Dataset

:::: logo-banner
<svg height="42" viewBox="0 0 272 92" xmlns="http://www.w3.org/2000/svg" aria-label="Google"> <path d="M115.75 47.18c0 12.77-9.99 22.18-22.25 22.18s-22.25-9.41-22.25-22.18C71.25 34.32 81.24 25 93.5 25s22.25 9.32 22.25 22.18zm-9.74 0c0-7.98-5.79-13.44-12.51-13.44S80.99 39.2 80.99 47.18c0 7.9 5.79 13.44 12.51 13.44s12.51-5.55 12.51-13.44z" fill="#EA4335"/> <path d="M163.75 47.18c0 12.77-9.99 22.18-22.25 22.18s-22.25-9.41-22.25-22.18c0-12.85 9.99-22.18 22.25-22.18s22.25 9.32 22.25 22.18zm-9.74 0c0-7.98-5.79-13.44-12.51-13.44s-12.51 5.46-12.51 13.44c0 7.9 5.79 13.44 12.51 13.44s12.51-5.55 12.51-13.44z" fill="#FBBC05"/> <path d="M209.75 26.34v39.82c0 16.38-9.66 23.07-21.08 23.07-10.75 0-17.22-7.19-19.66-13.07l8.48-3.53c1.51 3.61 5.21 7.87 11.17 7.87 7.31 0 11.84-4.51 11.84-13v-3.19h-.34c-2.18 2.69-6.38 5.04-11.68 5.04-11.09 0-21.25-9.66-21.25-22.09 0-12.52 10.16-22.26 21.25-22.26 5.29 0 9.49 2.35 11.68 4.96h.34v-3.61h9.25zm-8.56 20.92c0-7.81-5.21-13.52-11.84-13.52-6.72 0-12.35 5.71-12.35 13.52 0 7.73 5.63 13.36 12.35 13.36 6.63 0 11.84-5.63 11.84-13.36z" fill="#4285F4"/> <path d="M225 3v65h-9.5V3h9.5z" fill="#34A853"/> <path d="M262.02 54.48l7.56 5.04c-2.44 3.61-8.32 9.83-18.48 9.83-12.6 0-22.01-9.74-22.01-22.18 0-13.19 9.49-22.18 20.92-22.18 11.51 0 17.14 9.16 18.98 14.11l1.01 2.52-29.65 12.28c2.27 4.45 5.8 6.72 10.75 6.72 4.96 0 8.4-2.44 10.92-6.14zm-23.27-7.98l19.82-8.23c-1.09-2.77-4.37-4.7-8.23-4.7-4.95 0-11.84 4.37-11.59 12.93z" fill="#EA4335"/> <path d="M35.29 41.41V32h31.91c.31 1.64.47 3.58.47 5.68 0 7.06-1.93 15.79-8.15 22.01-6.05 6.3-13.78 9.66-24.02 9.66C16.32 69.35.36 53.89.36 34.67.36 15.45 16.32 0 35.5 0c10.32 0 17.22 4.03 22.6 9.16l-6.38 6.38c-3.86-3.61-9.09-6.38-16.22-6.38-13.27 0-23.65 10.67-23.65 24.11 0 13.44 10.38 24.11 23.65 24.11 8.6 0 13.52-3.44 16.64-6.56 2.54-2.54 4.21-6.17 4.87-11.12l-21.72.71z" fill="#4285F4"/> </svg>

::: logo-sep
:::

<svg height="42" viewBox="0 0 220 60" xmlns="http://www.w3.org/2000/svg" aria-label="Stanford SNAP"> <text y="28" font-family="Georgia, serif" font-size="22" font-weight="700" fill="#B83A3A" letter-spacing="1">STANFORD</text> <text y="52" font-family="Georgia, serif" font-size="16" font-weight="400" fill="#8c8c8c" letter-spacing="3">SNAP</text> </svg>
::::

-   Released by Google in $2002$ as part of the Google Programming Contest
-   Made publicly available through the Stanford Network Analysis Project (SNAP)
-   2002: aftermath of the dotcom bubble effecting the internet space

## Dataset Description

```{r load-data, include=FALSE}
library(igraph)
edges <- read.table("web-Google.txt.gz", comment.char = "#")
graph <- graph_from_data_frame(edges, directed = TRUE)
```

```{r, echo=FALSE,fig.align='center'}
avg.degree= mean(degree(graph))
max.degree = max(degree(graph))
density = edge_density(graph)

Dataset.info= data.frame(
  Property = c(
    "Nodes", "Edges", "Directionality", "Weighted", "Self-loops", " Density",
    "Mean Degree","Max Degree"
  ),
  Value = c(
    vcount(graph),
    ecount(graph),
    ifelse(is_directed(graph), "Yes","No"),
    ifelse(is_weighted(graph), "Yes","No"),
    ifelse(any(which_loop(graph)), "Yes","No"),
    round(density,7),
    round(avg.degree,2),
    max.degree
  )
)
knitr::kable(Dataset.info,
             caption = "Basic Network Elements of the web-Google Network",
             align = c("l","l"))
```

## `induced_subgraph()` Implemenation

```{r, echo=FALSE}
set.seed(123)

sub.network = lapply(1:4, function(x){
  induced_subgraph(graph, sample(V(graph),200))
})
sub.network[[1]]
sub.network[[2]]
sub.network[[3]]
sub.network[[4]]
```

## Random Walk Sampling Code

```{r}
set.seed(123)

subgraphs <- list()

for(i in 1:4){
  start_node <- sample(V(graph), 1)
  walk <- random_walk(graph,
                      start = start_node,
                      steps = 500,
                      mode = "all")
  nodes_sampled <- unique(walk)
  nodes_sampled <- nodes_sampled[1:200]
  subgraphs[[i]] <- induced_subgraph(graph, nodes_sampled)
}

```

## Checking Random Walk Sampling

```{r,echo=FALSE,fig.align='center'}
deg.full = degree(graph)

bias.results = data.frame(
  Subgraph = character(),
  Nodes = numeric(),
  Edges = numeric(),
  Mean_Degree = numeric(),
  Median_Degree = numeric(),
  KS_p_value = numeric(),
  stringsAsFactors = FALSE
)

for (i in 1:4){
  deg.sample = degree(subgraphs[[i]])
  ks = ks.test(deg.sample, deg.full)
  bias.results = rbind(
    bias.results,
    data.frame(
      Subgraph = paste("Subgraph", i),
      Nodes = vcount(subgraphs[[i]]),
      Edges = ecount(subgraphs[[i]]),
      Mean_Degree = mean(deg.sample),
      Median_Degree = median(deg.sample),
      KS_p_value = ks$p.value
      )
    )
}
knitr::kable(bias.results,
             caption = "Sampling Bias Assessment of Random Walk Subgraphs",
             col.names = c("Subgraph", "Nodes", "Edges", 
                           "Mean Degree", "Median Degree", "KS P-Value"),
             align = c("l", "r", "r", "r", "r", "r"))
```

## Subgraph 4 Visuals

```{r,echo=FALSE,fig.align='center'}
par(mfrow = c(1,3))

plot(
  subgraphs[[4]],
  layout = layout_with_fr(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - Fruchterman-Reingold"
)

plot(
  subgraphs[[4]],
  layout = layout_with_kk(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - Kamada-Kawai"
)

plot(
  subgraphs[[4]],
  layout = layout_with_mds(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - MDS"
)
```

```{r,echo=FALSE}
undirected_subgraphs <- lapply(subgraphs, function(g) as.undirected(g, mode = "collapse"))
```

## Metrics - Structure & Connectivity

```{r connectivity, echo=FALSE}
g <- subgraphs[[4]]
density    <- round(edge_density(g), 4)
strong_comp <- components(g, mode = "strong")$no
weak_comp   <- components(g, mode = "weak")$no
avg_path   <- round(mean(distances(g)[is.finite(distances(g))]), 3)
recip      <- round(reciprocity(g), 3)
trans      <- round(transitivity(g), 3)

metrics <- data.frame(
  Metric = c("Density", "Strong Components", "Weak Components",
             "Avg Path Length", "Reciprocity", "Transitivity"),
  Value = c(density, strong_comp, weak_comp, avg_path, recip, trans)
)

knitr::kable(metrics, align = c("l", "r"))
```

-   **Low density** confirms sparse web graph structure
-   **Reciprocity** reflects one-way nature of hyperlinks
-   **Transitivity** indicates degree of local clustering

## Community Detection

- Fast Greedy, Louvain, and Walktrap

```{r, echo=FALSE}
# list to store all results
all_results <- list()
g <- undirected_subgraphs[[4]]

objs <- list(
  fast_greedy = cluster_fast_greedy(g),
  louvain     = cluster_louvain(g),
  walktrap    = cluster_walktrap(g)
)

results <- data.frame(
  Subgraph      = i,
  Algorithm     = c("Fast Greedy", "Louvain", "Walktrap"),
  N_communities = c(length(objs$fast_greedy), length(objs$louvain), length(objs$walktrap)),
  Modularity_Q  = round(c(modularity(objs$fast_greedy), modularity(objs$louvain), modularity(objs$walktrap)), 4),
  Max_Comm_Size = c(max(sizes(objs$fast_greedy)), max(sizes(objs$louvain)), max(sizes(objs$walktrap)))
)

knitr::kable(results, align = c("l", "r"))
```

## Community Detection

- Walktrap detects more communities than other algorithms
  - Fast greedy and Louvain typically struggle to see smaller communities
  - Implies a hierarchical/subcommunity structure
- Stable results with modularity of around 0.7 or higher
- Largest community is less than a quarter of entire network size

## Adjacency Matrix

-   200x200 matrix where each cell = 1 if a hyperlink exists between two pages
-   Original ordering: edges appear scattered, no visible structure
-   Goal: reorder nodes to reveal hidden patterns

## Adjacency Matrix: Reordered by Community

-   Nodes reordered by Louvain community membership
-   Diagonal blocks emerge — pages in the same community densely hyperlink each other
-   Subgraph 4 (1,608 edges) shows the clearest block structure
-   Off-diagonal sparsity confirms pages from different communities rarely link directly

## Adjacency Matrix Reordered

```{r, fig.align='center', echo=FALSE}
adj_mat <- as_adjacency_matrix(subgraphs[[4]], sparse = FALSE)
kc_louv <- cluster_louvain(undirected_subgraphs[[4]])
node_order <- order(membership(kc_louv))
adj_mat_reordered <- adj_mat[node_order, node_order]
image(adj_mat_reordered,
      main = "Subgraph 4 - Reordered by Community",
      sub = "Blue = hyperlink exists, White = no hyperlink",
      col = c("white", "steelblue"),
      axes = FALSE,
      xlab = "Node (reordered)",
      ylab = "Node (reordered)")
```

## Reordering Nodes by Connected Component
- All 200 nodes in each subgraph belong to a single connected component
- Reordering by component yields identical results to community reordering
- Two patterns visible in reordered matrices:
  - **L-shape**: central hub page linking to many communities (e.g. Wikipedia-like pages)
  - **Solid block**: tightly interconnected pages, likely subpages of the same site


## Conclusion 

- Random Walk sampling allowed us to analyze meaningful locally connected regions for our sparse Google web graph

- Our sampled network consistently showed strong community structure, where webpages cluster into tightly connected groups

- High reciprocity and transitivity indicatet that webpages frequently link within their own communities

- Community detection algorithms produces high modularity ($\approx 0.7+$), confirming well-defined communities

- Walktrap detected more communities than Louvian and Fast-Greedy - hierarchical subcommunities within larger clusters

- Reordered adjacency matrices revealed clear block-diagonal patterns - hyperlinks are concentrated within topical clusters rather than randomly distributed

- Occasional hub-like pages connect multiple communities - bridges between different topics



## Future Direction

- Analyze larger sampled subgraphs to capture more global network structure
- Compare different network sampling methods to reduce sampling bias
- Explore temporal web datasets to study how web structure evolved over time



# Thank You






