---
title: "PSTAT194CS_SNA_Final_Project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,   # hides code
  message = FALSE,
  warning = FALSE
)
```

# Dataset

# Loading and Preparing Data
```{r}
url = "https://snap.stanford.edu/data/web-Google.txt.gz"
download.file(url, destfile = "web-Google.txt.gz", mode = "wb")
```
```{r}
library(igraph)

edges = read.table(
  "web-Google.txt.gz",
  comment.char = "#",
)

graph = graph_from_data_frame(edges, directed=TRUE)
```

# Understanding our Network
```{r}
cat("Number of Nodes:",vcount(graph),"\n")
cat("Number of Edges:",ecount(graph))
```

```{r}
is_directed(graph)
```



```{r}
is_weighted(graph)
```


```{r}
summary(graph)
```






# Creating Subgraphs of 200-nodes



## Code Implemenation
```{r}
set.seed(123)

sub.network = lapply(1:4, function(x){
  induced_subgraph(graph, sample(V(graph),200))
})
```

## Checking Subgraphs
```{r}
sub.network[[1]]
sub.network[[2]]
sub.network[[3]]
sub.network[[4]]
```










## Random Walk New Implementation
```{r}
set.seed(123)

subgraphs <- list()

for(i in 1:4){
  start_node <- sample(V(graph), 1)
  walk <- random_walk(graph,
                      start = start_node,
                      steps = 500,
                      mode = "all")
  nodes_sampled <- unique(walk)
  nodes_sampled <- nodes_sampled[1:200]
  subgraphs[[i]] <- induced_subgraph(graph, nodes_sampled)
}
```

## Checking our Subgraphs
```{r}
subgraphs[[1]]
subgraphs[[2]]
subgraphs[[3]]
subgraphs[[4]]
```


## Checking on Sampling Bias
```{r}
deg.full = degree(graph)


bias.results = data.frame(
  Subgraph = character(),
  Nodes = numeric(),
  Edges = numeric(),
  Mean_Degree = numeric(),
  Median_Degree = numeric(),
  KS_p_value = numeric(),
  stringsAsFactors = FALSE
)

for (i in 1:4){
  deg.sample = degree(subgraphs[[i]])
  ks = ks.test(deg.sample, deg.full)
  bias.results = rbind(
    bias.results,
    data.frame(
      Subgraph = paste("Subgraph", i),
      Nodes = vcount(subgraphs[[i]]),
      Edges = ecount(subgraphs[[i]]),
      Mean_Degree = mean(deg.sample),
      Median_Degree = median(deg.sample),
      KS_p_value = ks$p.value
      )
    )
}
bias.results
```













# Visualization 

## Subgraph 1
```{r}
par(mfrow = c(1,3))

plot(
  subgraphs[[1]],
  layout = layout_with_fr(subgraphs[[1]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 1 - Fruchterman-Reingold"
)

plot(
  subgraphs[[1]],
  layout = layout_with_kk(subgraphs[[1]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 1 - Kamada-Kawai"
)

plot(
  subgraphs[[1]],
  layout = layout_with_mds(subgraphs[[1]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 1 - MDS"
)
```

## Subgraph 2
```{r}
par(mfrow = c(1,3))

plot(
  subgraphs[[2]],
  layout = layout_with_fr(subgraphs[[2]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 2- Fruchterman-Reingold"
)

plot(
  subgraphs[[2]],
  layout = layout_with_kk(subgraphs[[2]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 2 - Kamada-Kawai"
)

plot(
  subgraphs[[2]],
  layout = layout_with_mds(subgraphs[[2]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 2 - MDS"
)
```



## Subgraph 3
```{r}
par(mfrow = c(1,3))

plot(
  subgraphs[[3]],
  layout = layout_with_fr(subgraphs[[3]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 3 - Fruchterman-Reingold"
)

plot(
  subgraphs[[3]],
  layout = layout_with_kk(subgraphs[[3]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 3 - Kamada-Kawai"
)

plot(
  subgraphs[[3]],
  layout = layout_with_mds(subgraphs[[3]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 3 - MDS"
)
```




## Subgraph 4
```{r}
par(mfrow = c(1,3))

plot(
  subgraphs[[4]],
  layout = layout_with_fr(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - Fruchterman-Reingold"
)

plot(
  subgraphs[[4]],
  layout = layout_with_kk(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - Kamada-Kawai"
)

plot(
  subgraphs[[4]],
  layout = layout_with_mds(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - MDS"
)
```













# Metrics Evaluated for ALL 4 SUBGRAPHS (probably don't do this)

```{r, echo = FALSE, eval = TRUE}
for (i in 1:4) {
  cat("Metrics for Subgraph", i, ":\n", "NETWORK LEVEL METRICS", "\n")
  
  cat("Density:", "\n") 
  print(edge_density(subgraphs[[i]]))

  # Reciprocity
  cat("Reciprocity:", "\n")
  print(reciprocity(subgraphs[[i]]))

  # Dyad Census 
  cat("Pairwise Relationships:", "\n")
  print(dyad_census(subgraphs[[i]]))

  # Transitivity 
  cat("Transitivity:", "\n")
  print(transitivity(subgraphs[[i]]))
  
  # Average distance between nodes
  cat("Mean Distance Between Nodes:", "\n")
  print(mean_distance(subgraphs[[i]]))

  # Diameter
  cat("Diameter and the Associated Nodes:", "\n")
  print(diameter(subgraphs[[i]], weights = NA))
  print(farthest_vertices(subgraphs[[i]]))
  
  # NODE LEVEL METRICS
  def_marg <- c(5.1, 4.1, 4.1, 2.1)
  cat("NODE LEVEL METRICS", "\n", "Degree:", "\n")
  print(degree(subgraphs[[i]], mode = "in"))
  par(mar= def_marg)
  hist(degree(subgraphs[[1]], mode = "in"),
       xlim = c(0, 40),
     breaks=1:vcount(subgraphs[[1]])-1, 
     main="Indegree Histogram")
  par(mar= def_marg)
  print(degree(subgraphs[[i]], mode = "out"))
  hist(degree(subgraphs[[1]], mode = "out"),
       xlim = c(0, 20),
     breaks=1:vcount(subgraphs[[1]])-1, 
     main="Outdegree Histogram")
  
  # Centrality
  cat("Betweenness Centrality:", "\n")
  print(betweenness(subgraphs[[i]]))
  cat(mean(is.nan(betweenness(subgraphs[[i]]))), "of the values are NaN", "\n")
  
  cat("Closeness Centrality:", "\n")
  clr <- centr_clo(subgraphs[[i]], mode="all", normalized=T)
  print(centr_clo(subgraphs[[i]], mode="all", normalized=T))
  cat(mean(is.nan(closeness(subgraphs[[i]]))), "of the values are NaN", "\n")
  sort(clr$res, decreasing = T, index.return = T)
  
  cat("Local Transitivity:", "\n")
  print(transitivity(subgraphs[[i]], type="local"))
}
```

## Interpretations

Density is the proportion of edges in a network compared to how many possible edges could exist. A density of `r edge_density(subgraphs[[1]])` indicates a sparse network with a low proportion of connected nodes. This is reasonable for a Google web graph, as only somewhat-related websites would have links connecting one to the other. 

Reciprocity is the proportion of edges from node A to B that have a corresponding edge traveling from B to A. This quantity tends to be higher in friendship networks where connections are mutual, and lower in follower networks due to their one-sided nature. Our value of `r reciprocity(subgraphs[[1]])` suggests a moderately low percentage of websites with mutual hyperlinks. 

The dyad census takes all pairs of nodes and calculates the number of mutual connections, one-sided (asymmetric) connections, and null connections among all node pairs. We see that there are `r dyad_census(subgraphs[[1]])$mut` mutually connected nodes and almost 20000 unconnected nodes. This echoes the low density found from earlier, and is perfectly reasonable since hyperlinks should be relevant to the website it was found on. 

Diameter measures the length of all shortest paths between nodes in the network and returns the maximum value, while average distance returns the mean. Higher values indicate more indirect connections. A mean distance of `r mean_distance(subgraphs[[1]])` is quite low, and suggests that connections seldom occur through a large number of intermediary websites.

Global transitivity is the likelihood that two different neighbors of a node also share an edge. High transitivity indicates a network with tight-knit groups and more clustering. This is the case for our network, as a transitivity of `r transitivity(subgraphs[[1]])` shows the close-knit nature of websites with hyperlinks conencting each other, even on a global scale. 

Degree is a node-level measure of how many edges connected to it. Since we are working with a directed graph, the number of in-edges and out-edges are distinct. There is a clear right skew in both histograms, which is expected for a web graph since most websites will have a handful of connections but few will have a great deal.

Betweenness is a centrality measure that finds the frequency of a node existing within the shortest path between two other nodes. Nodes with high betweenness centrality are often critical points in the network that control information flow. In our network some nodes have a betweenness measure of over 1400 while others have a betweenness of 0. Pages like Yahoo or trusted encyclopedia pages likely had these high centralities, since they would have been referred to frequently by other sites even if they were not directly related.  

Closeness is a centrality measure that measures the distance from a node to all other nodes. Important nodes will have a higher closeness since they interact with the other nodes more closely. The closeness values typically fall between 0.15 and 0.25, meaning that websites take around 4-7 hyperlinks to jump to the next website. The theme of close-knittedness between websites persists here; website hyperlinks will typically be relevant to the original website and most connections will not stray too far away. The low centralization of <0.15 suggests that centrality is distributed fairly evenly across the network, with very few websites dominating traffic.   

Local transitivity is the smaller-scale measure of transitivity, and tells whether the data's clustering occurs frequently between small groups. Values frequently take values close to 1, which again shows that web graphs are tightly connected on a local level. Sites covering related topics will likely reference each other. 


  
  

# 5) Community Detection

Because the community detection algorithms want an undirected graph and our current graph is directed, we need to convert it into an undirected graph first. Then, we can run all of the community detection algorithms on each subgraph and store it all inside one dataframe. We will also store the results inside a comunity objects data structure so that we can use the results later.

```{r}
# convert all into undirected graphs
undirected_subgraphs <- vector("list", 4)
for (i in 1:4){
  undirected_subgraphs[[i]] <- as.undirected(subgraphs[[i]], mode="collapse")
}
# list to store all results
all_results <- list()
community_objects <- list()

for (i in 1:4) {
  g <- undirected_subgraphs[[i]]
  
  community_objects[[i]] <- list(
    fast_greedy = cluster_fast_greedy(g),
    louvain     = cluster_louvain(g),
    walktrap    = cluster_walktrap(g)
  )
  
  # Grab the objects we just saved to calculate stats
  objs <- community_objects[[i]]
  
  all_results[[i]] <- data.frame(
    Subgraph      = i,
    Algorithm     = c("Fast Greedy", "Louvain", "Walktrap"),
    N_communities = c(length(objs$fast_greedy), length(objs$louvain), length(objs$walktrap)),
    Modularity_Q  = round(c(modularity(objs$fast_greedy), modularity(objs$louvain), modularity(objs$walktrap)), 4),
    Max_Comm_Size = c(max(sizes(objs$fast_greedy)), max(sizes(objs$louvain)), max(sizes(objs$walktrap)))
  )
}

# combine list into one table
final_comparison <- do.call(rbind, all_results)
final_comparison
```

Looking at the results, we see that Walktrap generally detects more communities than the other 2 algorithms. All of the algorithms have high modularity greater than 0.7, suggesting that they found clear communities.


# 6) Adjacency Matrix
```{r}
for(i in 1:4){
  adj_mat <- as_adjacency_matrix(subgraphs[[i]], sparse = FALSE)
  
  image(adj_mat,
        main = paste("Subgraph", i, "- Original Node Order"),
        col = c("white", "steelblue"),
        axes = FALSE)
}
# convert all into undirected graphs
undirected_subgraphs <- vector("list", 4)
for (i in 1:4){
  undirected_subgraphs[[i]] <- as.undirected(subgraphs[[i]])
}
# list to store all results
all_results <- list()

for (i in 1:4) {
  g <- undirected_subgraphs[[i]]
  
  # run algs
  fg   <- cluster_fast_greedy(g)
  louv <- cluster_louvain(g)
  walk <- cluster_walktrap(g)
  
  # store into list
  all_results[[i]] <- data.frame(
    Subgraph      = i,
    Algorithm     = c("Fast Greedy", "Louvain", "Walktrap"),
    N_communities = c(length(fg), length(louv), length(walk)),
    Modularity_Q  = round(c(modularity(fg), modularity(louv), modularity(walk)), 4),
    Max_Comm_Size = c(max(sizes(fg)), max(sizes(louv)), max(sizes(walk)))
  )
}

# combine list into one table
final_comparison <- do.call(rbind, all_results)
final_comparison
```



# 6) Adjacency Matrix
```{r}
for(i in 1:4){
  adj_mat <- as_adjacency_matrix(subgraphs[[i]], sparse = FALSE)
  
  image(adj_mat,
        main = paste("Subgraph", i, "- Original Node Order"),
        col = c("white", "steelblue"),
        axes = FALSE)
}
```


```{r}
set.seed(42)
for(i in 1:4){
  adj_mat <- as_adjacency_matrix(subgraphs[[i]], sparse = FALSE)
  
  kc_louv <- cluster_louvain(undirected_subgraphs[[i]])
  node_order <- order(membership(kc_louv))
  adj_mat_reordered <- adj_mat[node_order, node_order]
  
  image(adj_mat_reordered,
        main = paste("Subgraph", i, "- Reordered by Community"),
        col = c("white", "steelblue"),
        axes = FALSE)
}
```


# 7) Reordering Nodes for Community Detection

As we see above, all of the nodes are in one connected component. As such, this step is essentially the same as above. We will reorder the adjacency matrices for all of the remaining subgraphs as well.

```{r}
for (i in 1:4) {
  A <- as_adjacency_matrix(undirected_subgraphs[[i]], sparse = FALSE)
  
  par(mfrow = c(1, 3))
  
  # loop through the algorithms
  alg_names <- names(community_objects[[i]])
  
  for (alg in alg_names) {
    comm_obj <- community_objects[[i]][[alg]]
    
    ord <- order(membership(comm_obj))
    
    A_reordered <- A[ord, ord]
    
    image(A_reordered,
          main = paste("Sub", i, ":", alg),
          xlab = "Node (reordered)", 
          ylab = "Node (reordered)",
          col=c("white", "steelblue"),
          axes = FALSE)
  }
  
  # reset plots for next one
  par(mfrow = c(1, 1))
}
```

After reordering by community for each algorithm, we can see that the nodes form multiple boxes. This represents communities where nodes are highly interconnected. The adjacency matrix also happens to be mirrored across the y=x line, and that's because we used the undirected graph. We chose to use the undirected graph instead of the directed graph because that is the graph the algorithms used to detect the communities. If you also compare the adjencency matrix to the algorithm results, you see that the boxes correspond to how many communities each algorithm detected.

There are two main shapes we see in the adjacency matrix. First is this sparse L shape. This represents a central node that connects to many other nodes. In the context of Google, this could be a site like Wikipedia that has links to many different communities. Next is a solid box. This represents a community where all the nodes are closely interconnected. This could be a set of subpages, where all pages link to each other. 





























\newpage

## Appendix

```{r appendix-code, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
