---
title: "PSTAT194CS_SNA_Final_Project"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,   # hides code
  message = FALSE,
  warning = FALSE
)
```

# Dataset

# Loading and Preparing Data
```{r}
url = "https://snap.stanford.edu/data/web-Google.txt.gz"
download.file(url, destfile = "web-Google.txt.gz", mode = "wb")
```
```{r}
library(igraph)

edges = read.table(
  "web-Google.txt.gz",
  comment.char = "#",
)

graph = graph_from_data_frame(edges, directed=TRUE)
```

# Understanding our Network
```{r}
cat("Number of Nodes:",vcount(graph),"\n")
cat("Number of Edges:",ecount(graph))
```

```{r}
is_directed(graph)
```



```{r}
is_weighted(graph)
```


```{r}
summary(graph)
```






# Creating Subgraphs of 200-nodes



## Code Implemenation
```{r, eval=FALSE}
set.seed(123)

sub.network = lapply(1:4, function(x){
  induced_subgraph(graph, sample(V(graph),200))
})
```

## Checking Subgraphs
```{r, eval=FALSE}
sub.network[[1]]
sub.network[[2]]
sub.network[[3]]
sub.network[[4]]
```










## Random Walk New Implementation
```{r}
set.seed(123)

subgraphs <- list()

for(i in 1:4){
  start_node <- sample(V(graph), 1)
  walk <- random_walk(graph,
                      start = start_node,
                      steps = 500,
                      mode = "all")
  nodes_sampled <- unique(walk)
  nodes_sampled <- nodes_sampled[1:200]
  subgraphs[[i]] <- induced_subgraph(graph, nodes_sampled)
}
```

## Checking our Subgraphs
```{r}
subgraphs[[1]]
subgraphs[[2]]
subgraphs[[3]]
subgraphs[[4]]
```


## Checking on Sampling Bias
```{r}
deg.full = degree(graph)


bias.results = data.frame(
  Subgraph = character(),
  Nodes = numeric(),
  Edges = numeric(),
  Mean_Degree = numeric(),
  Median_Degree = numeric(),
  KS_p_value = numeric(),
  stringsAsFactors = FALSE
)

for (i in 1:4){
  deg.sample = degree(subgraphs[[i]])
  ks = ks.test(deg.sample, deg.full)
  bias.results = rbind(
    bias.results,
    data.frame(
      Subgraph = paste("Subgraph", i),
      Nodes = vcount(subgraphs[[i]]),
      Edges = ecount(subgraphs[[i]]),
      Mean_Degree = mean(deg.sample),
      Median_Degree = median(deg.sample),
      KS_p_value = ks$p.value
      )
    )
}
bias.results
```













# Visualization 

```{r}
layouts = list(layout_with_fr, layout_with_kk, layout_with_mds)

for(i in 1:4){
  for (j in 1:3){
    plot(
      subgraphs[[i]],
      layout = layouts[[j]],
      vertex.size = 4,
      vertex.label = NA,
      edge.arrow.size = 0.2,
      main= paste("Subgraph", i,"- Layout", j)
    )
  }
}
```

# Metrics Evaluated for ALL 4 SUBGRAPHS (probably don't do this)

```{r, eval = TRUE, results='hide'}
for (i in 1:4) {
  cat("Metrics for Subgraph", i, ":\n", "NETWORK LEVEL METRICS", "\n")
  
  cat("Density:", "\n") 
  print(edge_density(subgraphs[[i]]))

  # Reciprocity
  cat("Reciprocity:", "\n")
  print(reciprocity(subgraphs[[i]]))

  # Dyad Census 
  cat("Pairwise Relationships:", "\n")
  print(dyad_census(subgraphs[[i]]))

  # Transitivity 
  cat("Transitivity:", "\n")
  print(transitivity(subgraphs[[i]]))
  
  # Average distance between nodes
  cat("Mean Distance Between Nodes:", "\n")
  print(mean_distance(subgraphs[[i]]))

  # Diameter
  cat("Diameter and the Associated Nodes:", "\n")
  print(diameter(subgraphs[[i]], weights = NA))
  print(farthest_vertices(subgraphs[[i]]))
  
  # NODE LEVEL METRICS
  cat("NODE LEVEL METRICS", "\n", "Degree:", "\n")
  print(degree(subgraphs[[i]], mode = "in"))
  
  # Centrality
  cat("Betweenness Centrality:", "\n")
  print(betweenness(subgraphs[[i]]))
  cat(mean(is.nan(betweenness(subgraphs[[i]]))), "of the values are NaN", "\n")
  
  cat("Closeness Centrality:", "\n")
  clr <- centr_clo(subgraphs[[i]], mode="all", normalized=T)
  print(centr_clo(subgraphs[[i]], mode="all", normalized=T))
  cat(mean(is.nan(closeness(subgraphs[[i]]))), "of the values are NaN", "\n")
  sort(clr$res, decreasing = T, index.return = T) # REMOVE ME Can remove if we don't want to plot
  
  cat("Local Transitivity:", "\n")
  print(transitivity(subgraphs[[i]], type="local"))

  # Edge Level
  edge_betweenness(subgraphs[[i]])
}
```

# 5) Community Detection

Because the community detection algorithms want an undirected graph and our current graph is directed, we need to convert it into an undirected graph first. Then, we can run all of the community detection algorithms on each subgraph and store it all inside one dataframe. We will also store the results inside a comunity objects data structure so that we can use the results later.

```{r}
# convert all into undirected graphs
undirected_subgraphs <- vector("list", 4)
for (i in 1:4){
  undirected_subgraphs[[i]] <- as.undirected(subgraphs[[i]], mode="collapse")
}
# list to store all results
all_results <- list()
community_objects <- list()

for (i in 1:4) {
  g <- undirected_subgraphs[[i]]
  
  community_objects[[i]] <- list(
    fast_greedy = cluster_fast_greedy(g),
    louvain     = cluster_louvain(g),
    walktrap    = cluster_walktrap(g)
  )
  
  # Grab the objects we just saved to calculate stats
  objs <- community_objects[[i]]
  
  all_results[[i]] <- data.frame(
    Subgraph      = i,
    Algorithm     = c("Fast Greedy", "Louvain", "Walktrap"),
    N_communities = c(length(objs$fast_greedy), length(objs$louvain), length(objs$walktrap)),
    Modularity_Q  = round(c(modularity(objs$fast_greedy), modularity(objs$louvain), modularity(objs$walktrap)), 4),
    Max_Comm_Size = c(max(sizes(objs$fast_greedy)), max(sizes(objs$louvain)), max(sizes(objs$walktrap)))
  )
}

# combine list into one table
final_comparison <- do.call(rbind, all_results)
final_comparison
```

Looking at the results, we see that Walktrap generally detects more communities than the other 2 algorithms. All of the algorithms have high modularity greater than 0.7, suggesting that they found clear communities.


# 6) Adjacency Matrix
```{r}
for(i in 1:4){
  adj_mat <- as_adjacency_matrix(subgraphs[[i]], sparse = FALSE)
  
  image(adj_mat,
        main = paste("Subgraph", i, "- Original Node Order"),
        col = c("white", "steelblue"),
        axes = FALSE)
}
```


```{r}
set.seed(42)
for(i in 1:4){
  adj_mat <- as_adjacency_matrix(subgraphs[[i]], sparse = FALSE)
  
  kc_louv <- cluster_louvain(undirected_subgraphs[[i]])
  node_order <- order(membership(kc_louv))
  adj_mat_reordered <- adj_mat[node_order, node_order]
  
  image(adj_mat_reordered,
        main = paste("Subgraph", i, "- Reordered by Community"),
        col = c("white", "steelblue"),
        axes = FALSE)
}
```

The adjacency matrix of each 200-node subgraph was examined to identify structural patterns in the Google web graph. Each entry in the matrix represents the presence of a directed hyperlink between two pages. In the original node ordering, edges appear scattered with no discernible pattern, reflecting the arbitrary order in which nodes were sampled.
To reveal underlying structure, nodes were reordered by Louvain community membership. After reordering, clear diagonal block patterns emerged, particularly in Subgraph 4 which contained 1,608 edges. Each block along the diagonal represents a group of web pages that densely hyperlink to one another, likely corresponding to topically related pages. The sparsity off the diagonal indicates that pages from different communities rarely link directly to each other. This visual confirmation aligns with the community detection results, supporting the presence of meaningful cluster structure within the Google web graph.

# 7) Reordering Nodes for Community Detection

As we see above, all of the nodes are in one connected component. As such, this step is essentially the same as above. We will reorder the adjacency matrices for all of the remaining subgraphs as well.

```{r}
for (i in 1:4) {
  A <- as_adjacency_matrix(undirected_subgraphs[[i]], sparse = FALSE)
  
  par(mfrow = c(1, 3))
  
  # loop through the algorithms
  alg_names <- names(community_objects[[i]])
  
  for (alg in alg_names) {
    comm_obj <- community_objects[[i]][[alg]]
    
    ord <- order(membership(comm_obj))
    
    A_reordered <- A[ord, ord]
    
    image(A_reordered,
          main = paste("Sub", i, ":", alg),
          xlab = "Node (reordered)", 
          ylab = "Node (reordered)",
          col=c("white", "steelblue"),
          axes = FALSE)
  }
  
  # reset plots for next one
  par(mfrow = c(1, 1))
}
```

After reordering by community for each algorithm, we can see that the nodes form multiple boxes. This represents communities where nodes are highly interconnected. The adjacency matrix also happens to be mirrored across the y=x line, and that's because we used the undirected graph. We chose to use the undirected graph instead of the directed graph because that is the graph the algorithms used to detect the communities. If you also compare the adjencency matrix to the algorithm results, you see that the boxes correspond to how many communities each algorithm detected.

There are two main shapes we see in the adjacency matrix. First is this sparse L shape. This represents a central node that connects to many other nodes. In the context of Google, this could be a site like Wikipedia that has links to many different communities. Next is a solid box. This represents a community where all the nodes are closely interconnected. This could be a set of subpages, where all pages link to each other. 





























\newpage

## Appendix

```{r appendix-code, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
