---
title: "A Random Walk Through the Web: Network Analysis of the Google Web Graph"
author: "Richard Zhang, Toby Liu, Frank Wang, Max Peng"
subtitle: PSTAT 194CS
output:
  slidy_presentation: default
  beamer_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(igraph)
```

## Google Dataset

-   Released by Google in $2002$ as part of the Google Programming Contest
-   Made publicly available through the Stanford Network Analysis Project (SNAP)

## Dataset Description

```{r load-data, include=FALSE}
library(igraph)
edges <- read.table("web-Google.txt.gz", comment.char = "#")
graph <- graph_from_data_frame(edges, directed = TRUE)
```

```{r, echo=FALSE,fig.align='center'}
avg.degree= mean(degree(graph))
max.degree = max(degree(graph))
density = edge_density(graph)

Dataset.info= data.frame(
  Property = c(
    "Nodes", "Edges", "Directionality", "Weighted", "Self-loops", " Density",
    "Mean Degree","Max Degree"
  ),
  Value = c(
    vcount(graph),
    ecount(graph),
    ifelse(is_directed(graph), "Yes","No"),
    ifelse(is_weighted(graph), "Yes","No"),
    ifelse(any(which_loop(graph)), "Yes","No"),
    round(density,7),
    round(avg.degree,2),
    max.degree
  )
)
knitr::kable(Dataset.info,
             caption = "Basic Network Elements of the web-Google Network",
             align = c("l","l"))
```

## `induced_subgraph()` Implemenation

```{r, echo=FALSE}
set.seed(123)

sub.network = lapply(1:4, function(x){
  induced_subgraph(graph, sample(V(graph),200))
})
sub.network[[1]]
sub.network[[2]]
sub.network[[3]]
sub.network[[4]]
```

## Random Walk Sampling Code

```{r}
set.seed(123)

subgraphs <- list()

for(i in 1:4){
  start_node <- sample(V(graph), 1)
  walk <- random_walk(graph,
                      start = start_node,
                      steps = 500,
                      mode = "all")
  nodes_sampled <- unique(walk)
  nodes_sampled <- nodes_sampled[1:200]
  subgraphs[[i]] <- induced_subgraph(graph, nodes_sampled)
}

```

## Checking Random Walk Sampling

```{r,echo=FALSE,fig.align='center'}
deg.full = degree(graph)

bias.results = data.frame(
  Subgraph = character(),
  Nodes = numeric(),
  Edges = numeric(),
  Mean_Degree = numeric(),
  Median_Degree = numeric(),
  KS_p_value = numeric(),
  stringsAsFactors = FALSE
)

for (i in 1:4){
  deg.sample = degree(subgraphs[[i]])
  ks = ks.test(deg.sample, deg.full)
  bias.results = rbind(
    bias.results,
    data.frame(
      Subgraph = paste("Subgraph", i),
      Nodes = vcount(subgraphs[[i]]),
      Edges = ecount(subgraphs[[i]]),
      Mean_Degree = mean(deg.sample),
      Median_Degree = median(deg.sample),
      KS_p_value = ks$p.value
      )
    )
}
knitr::kable(bias.results,
             caption = "Sampling Bias Assessment of Random Walk Subgraphs",
             col.names = c("Subgraph", "Nodes", "Edges", 
                           "Mean Degree", "Median Degree", "KS P-Value"),
             align = c("l", "r", "r", "r", "r", "r"))
```

## Subgraph 4 Visuals

```{r,echo=FALSE,fig.align='center'}
par(mfrow = c(1,3))

plot(
  subgraphs[[4]],
  layout = layout_with_fr(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - Fruchterman-Reingold"
)

plot(
  subgraphs[[4]],
  layout = layout_with_kk(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - Kamada-Kawai"
)

plot(
  subgraphs[[4]],
  layout = layout_with_mds(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - MDS"
)
```

```{r,echo=FALSE}
undirected_subgraphs <- lapply(subgraphs, function(g) as.undirected(g, mode = "collapse"))
```

## Metrics - Structure & Connectivity

```{r connectivity, echo=FALSE}
g <- subgraphs[[4]]
density    <- round(edge_density(g), 4)
strong_comp <- components(g, mode = "strong")$no
weak_comp   <- components(g, mode = "weak")$no
avg_path   <- round(mean(distances(g)[is.finite(distances(g))]), 3)
recip      <- round(reciprocity(g), 3)
trans      <- round(transitivity(g), 3)

metrics <- data.frame(
  Metric = c("Density", "Strong Components", "Weak Components",
             "Avg Path Length", "Reciprocity", "Transitivity"),
  Value = c(density, strong_comp, weak_comp, avg_path, recip, trans)
)

knitr::kable(metrics, align = c("l", "r"))
```

-   **Low density** confirms sparse web graph structure
-   **Reciprocity** reflects one-way nature of hyperlinks
-   **Transitivity** indicates degree of local clustering

## Community Detection

- Fast Greedy, Louvain, and Walktrap

```{r, echo=FALSE}
# list to store all results
all_results <- list()
g <- undirected_subgraphs[[4]]

community_objects[[i]] <- list(
  fast_greedy = cluster_fast_greedy(g),
  louvain     = cluster_louvain(g),
  walktrap    = cluster_walktrap(g)
)

# Grab the objects we just saved to calculate stats
objs <- community_objects[[i]]

all_results[[i]] <- data.frame(
  Subgraph      = i,
  Algorithm     = c("Fast Greedy", "Louvain", "Walktrap"),
  N_communities = c(length(objs$fast_greedy), length(objs$louvain), length(objs$walktrap)),
  Modularity_Q  = round(c(modularity(objs$fast_greedy), modularity(objs$louvain), modularity(objs$walktrap)), 4),
  Max_Comm_Size = c(max(sizes(objs$fast_greedy)), max(sizes(objs$louvain)), max(sizes(objs$walktrap)))
)

```

## Community Detection

- Walktrap detects more communities than other algorithms
- Implies a hierarchical/subcommunity structure
- Stable results with modularity of around 0.7 or higher

## Adjacency Matrix

-   200x200 matrix where each cell = 1 if a hyperlink exists between two pages
-   Original ordering: edges appear scattered, no visible structure
-   Goal: reorder nodes to reveal hidden patterns

## Adjacency Matrix: Reordered by Community

-   Nodes reordered by Louvain community membership
-   Diagonal blocks emerge — pages in the same community densely hyperlink each other
-   Subgraph 4 (1,608 edges) shows the clearest block structure
-   Off-diagonal sparsity confirms pages from different communities rarely link directly

## Adjacency Matrix Reordered

```{r, fig.align='center', echo=FALSE}
adj_mat <- as_adjacency_matrix(subgraphs[[4]], sparse = FALSE)
kc_louv <- cluster_louvain(undirected_subgraphs[[4]])
node_order <- order(membership(kc_louv))
adj_mat_reordered <- adj_mat[node_order, node_order]
image(adj_mat_reordered,
      col = c("white", "steelblue"),
      axes = FALSE)
```
