---
title: "PSTAT194CS_SNA_Final_Project"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,   # hides code
  message = FALSE,
  warning = FALSE
)
```
\begin{titlepage}
  \centering
  \vspace*{2cm}
  
  {\LARGE \textbf{A Random Walk Through the Web}\par}
  {\LARGE \textbf{Network Analysis of the Google Web Graph}\par}
  
  \vspace{1.5cm}
  
  {\large Richard Zhang, Max Peng, Toby Liu, Frank Wang\par}
  
  \vspace{0.5cm}
  
  {\normalsize PSTAT 194CS — Social Network Analysis\par}
  {\normalsize University of California, Santa Barbara\par}
  
  \vspace{0.5cm}
  
  {\normalsize March 2026\par}
  
  \vfill
  
\end{titlepage}
# Introduction
The internet has become one of the most transformative and essential technologies in the early AI era. Serving as a universal learning tool, research platform, and the foundational infrastructure where AI continues to advance. With the rapid rise of AI, the internet has never been more critical, powering data pipelines, knowledge bases, and interconnected systems that modern machine learning systems depend on. Yet to truly understand the internet as it is today, it is valuable to examine it in its early development. Particularly in $2002$, during a turbulent times for the internet as it marked the aftermath of the dotcom bubble. Which led to many internet-based companies to rise and fall almost instantly. Fundamentally reshaping how the internet is today. 

Understanding how web pages were interconnected during this era not only reveals the structural properties of the early internet but also sheds light on how information flowed, how communities of related pages formed, and how search engines like Google first learned to navigate and rank content in the wake of one of the most dramatic technological disruptions in modern history.

## Dataset Description

We will be conducting a social network analysis to the web-Google dataset. Which is a directed web graph released by Google in $2002$ as part of the Google Programming Contest. Made publicly available through the Stanford Network Analysis Project (SNAP). Where the dataset comprises of $875713$ webpages (vertices) and $5105039$ directed hyperlinks (edges) capturing a rich and large-scale portait of the internet as it existed in $2002$. In which, the vertices represent individual web pages that were collected from the internet in $2002$. With each node corresponding to a unique webpage labeled by a numeric id, and edges representing directed hyperlinks between webpages.


```{r, echo=FALSE}
#Loading and Preparing Dataset
library(igraph)

url = "https://snap.stanford.edu/data/web-Google.txt.gz"
download.file(url, destfile = "web-Google.txt.gz", mode = "wb")

edges = read.table(
  "web-Google.txt.gz",
  comment.char = "#",
)

graph = graph_from_data_frame(edges, directed=TRUE)
```

```{r, echo=FALSE}
avg.degree= mean(degree(graph))
max.degree = max(degree(graph))
density = edge_density(graph)

Dataset.info= data.frame(
  Property = c(
    "Nodes", "Edges", "Directionality", "Weighted", "Self-loops", " Density",
    "Mean Degree","Max Degree"
  ),
  Value = c(
    vcount(graph),
    ecount(graph),
    ifelse(is_directed(graph), "Yes","No"),
    ifelse(is_weighted(graph), "Yes","No"),
    ifelse(any(which_loop(graph)), "Yes","No"),
    round(density,7),
    round(avg.degree,2),
    max.degree
  )
)
knitr::kable(Dataset.info,
             caption = "Basic Network Elements of the web-Google Network",
             align = c("l","l"))
```






## Objectives:
The objectives of this analysis are to:
* Generate representative subgraphs of the Google Web graph using random walk sampling to deal with sparsity limitations of induced subgraphs
* Look into sampling bias by comparing degree distributions between the sampled subgraph and full network using the Kolmpgorov-Smirnov test
* Visualize the network structure of our subgraphs with the Fruchterman-Reingold, Kamada-Kawai, and Multidimensional Scaling (MDS) layout to reveal structural and connectivity patterns
* Compute and interpret network-level metrics (i.e. density, reciprocity, dyad census, transitivity, diameter, and mean distance) to capture the overall structure of the web graph 
* Compute and interpret node-level metrics (i.e. in-degree, out-degree, betweenness centrality, closness centrality, and local transitivity) to identify influential web pages within the network
* Detect and compare communities using Fast-Greedy, Louvain, Walktrap algorithms and evaluate their results using molecularity scores and community sizes
* Examine the adjacency matrix of each subgraphs in their original and reordered form to confirm the presence of community strucutre and identify structural patterns.

# Methodology

## Creating Subgraphs

We will first generate four induced subgraphs by "purely" randomly sample $200$ nodes from the full network just using the `induced_subgraph()`. 
```{r, echo=FALSE}
set.seed(123)

sub.network = lapply(1:4, function(x){
  induced_subgraph(graph, sample(V(graph),200))
})
sub.network[[1]]
sub.network[[2]]
sub.network[[3]]
sub.network[[4]]
```
However, we can notice that from the results above, our subgraphs virtually do not have any edges. This could be due to the fact that the web-Google graph is extremely sparse (density $\approx$ `r round(edge_density(graph),7)`). This would make it difficult to get any meaningful analysis. 

To resolve this issue, we will utilize the concept of random walk to sample. Where we would use igraph's built in `random_walk()` function. Beginning with a randomly chosen seed node, a random walk proceeds for $500$ steps by following the network’s edges and recording the nodes encountered. The first $200$ unique nodes visited during this process are then used to construct each induced subgraph. This method helps ensure that the resulting subgraphs are well connected and contain many edges, maintaining the local hyperlink structure of the web.

### Code Implementation
```{r}
set.seed(123)

subgraphs <- list()

for(i in 1:4){
  start_node <- sample(V(graph), 1)
  walk <- random_walk(graph,
                      start = start_node,
                      steps = 500,
                      mode = "all")
  nodes_sampled <- unique(walk)
  nodes_sampled <- nodes_sampled[1:200]
  subgraphs[[i]] <- induced_subgraph(graph, nodes_sampled)
}


```

```{r,echo=FALSE}

subgraph.info = data.frame(
  Subgraph = paste("Subgraph", 1:4),
  Nodes = sapply(subgraphs, vcount),
  Edges = sapply(subgraphs, ecount)
)

knitr::kable(subgraph.info,
             caption = "Summary of Random Walk Subgraphs",
             align = c("l","r","r"))
```
Based on the results above, we can see that our random walk sampling helped with providing us with edges for our all four of our subgraphs.

## Checking on Sampling Bias

Though our subgraphs have edges using random walk sampling, there are still sampling biases that we need to consider. 
```{r}
deg.full = degree(graph)

bias.results = data.frame(
  Subgraph = character(),
  Nodes = numeric(),
  Edges = numeric(),
  Mean_Degree = numeric(),
  Median_Degree = numeric(),
  KS_p_value = numeric(),
  stringsAsFactors = FALSE
)

for (i in 1:4){
  deg.sample = degree(subgraphs[[i]])
  ks = ks.test(deg.sample, deg.full)
  bias.results = rbind(
    bias.results,
    data.frame(
      Subgraph = paste("Subgraph", i),
      Nodes = vcount(subgraphs[[i]]),
      Edges = ecount(subgraphs[[i]]),
      Mean_Degree = mean(deg.sample),
      Median_Degree = median(deg.sample),
      KS_p_value = ks$p.value
      )
    )
}
knitr::kable(bias.results,
             caption = "Sampling Bias Assessment of Random Walk Subgraphs",
             col.names = c("Subgraph", "Nodes", "Edges", 
                           "Mean Degree", "Median Degree", "KS P-Value"),
             align = c("l", "r", "r", "r", "r", "r"))
```
By comparing the degree distributions of each subgraph against the full network using summary statistics and the Kolmogorov-Smirnov (KS) test. Looking into our mean degree across subgraphs, we can notice that they range from $8.42$ to $16.08$. Where our median degree range from $5$ to $12$. Which suggests that some walks landed in denser regions of the network than others. Particularly subgraph $4$. As it is considerably denser than the other subgraphs. Indicating that the random walk from subgraph $4$ traversed a more tightly connected neighborhood of the web graph. This is important as we will be utilizing subgraph $4$ to observe network metrics later in our report. 

When observing our KS test p-values, we can see that all four subgraphs result in extremely small p-values. Indicating that a statistically significant difference between the subgraphs and the full network degree distributions. This is not bad as this is expected given the massive size difference between $200$ nodes and the full roughly $876000$ nodes. With this scale the KS test is extremely sensitive. Flagging even the smallest differences as significant. Thus our random walk sampled subgraphs are still suitable for structural analysis. As it still preserves the local connectivity patterns of the web graph rather than attempting to replocate irs global degree distribution. 

# Analysis

## Visualization 

### Subgraph 1
```{r, echo=FALSE}
par(mfrow = c(1,3))

plot(
  subgraphs[[1]],
  layout = layout_with_fr(subgraphs[[1]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 1 - Fruchterman-Reingold"
)

plot(
  subgraphs[[1]],
  layout = layout_with_kk(subgraphs[[1]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 1 - Kamada-Kawai"
)

plot(
  subgraphs[[1]],
  layout = layout_with_mds(subgraphs[[1]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 1 - MDS"
)
```

### Subgraph 2
```{r,echo=FALSE}
par(mfrow = c(1,3))

plot(
  subgraphs[[2]],
  layout = layout_with_fr(subgraphs[[2]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 2- Fruchterman-Reingold"
)

plot(
  subgraphs[[2]],
  layout = layout_with_kk(subgraphs[[2]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 2 - Kamada-Kawai"
)

plot(
  subgraphs[[2]],
  layout = layout_with_mds(subgraphs[[2]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 2 - MDS"
)
```



### Subgraph 3
```{r,echo=FALSE}
par(mfrow = c(1,3))

plot(
  subgraphs[[3]],
  layout = layout_with_fr(subgraphs[[3]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 3 - Fruchterman-Reingold"
)

plot(
  subgraphs[[3]],
  layout = layout_with_kk(subgraphs[[3]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 3 - Kamada-Kawai"
)

plot(
  subgraphs[[3]],
  layout = layout_with_mds(subgraphs[[3]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 3 - MDS"
)
```




### Subgraph 4
```{r,echo=FALSE}
par(mfrow = c(1,3))

plot(
  subgraphs[[4]],
  layout = layout_with_fr(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - Fruchterman-Reingold"
)

plot(
  subgraphs[[4]],
  layout = layout_with_kk(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - Kamada-Kawai"
)

plot(
  subgraphs[[4]],
  layout = layout_with_mds(subgraphs[[4]]),
  vertex.size = 4,
  vertex.label = NA,
  edge.arrow.size = 0.2,
  main = "Subgraph 4 - MDS"
)
```



## Network Metrics

```{r, echo = FALSE, eval = TRUE}
cat("Metrics for Subgraph", i, ":\n", "NETWORK LEVEL METRICS", "\n")

cat("Density:", "\n") 
print(edge_density(subgraphs[[4]]))

# Reciprocity
cat("Reciprocity:", "\n")
print(reciprocity(subgraphs[[4]]))

# Dyad Census 
cat("Pairwise Relationships:", "\n")
print(dyad_census(subgraphs[[4]]))

# Transitivity 
cat("Transitivity:", "\n")
print(transitivity(subgraphs[[4]]))

# Average distance between nodes
cat("Mean Distance Between Nodes:", "\n")
print(mean_distance(subgraphs[[4]]))

# Diameter
cat("Diameter and the Associated Nodes:", "\n")
print(diameter(subgraphs[[4]], weights = NA))
print(farthest_vertices(subgraphs[[4]]))

# NODE LEVEL METRICS
def_marg <- c(5.1, 4.1, 4.1, 2.1)
cat("NODE LEVEL METRICS", "\n", "Degree:", "\n")
print(degree(subgraphs[[4]], mode = "in"))
par(mar= def_marg)
hist(degree(subgraphs[[4]], mode = "in"),
     xlim = c(0, 40),
   breaks=1:vcount(subgraphs[[4]])-1, 
   main="Indegree Histogram")
par(mar= def_marg)
print(degree(subgraphs[[4]], mode = "out"))
hist(degree(subgraphs[[4]], mode = "out"),
     xlim = c(0, 20),
   breaks=1:vcount(subgraphs[[4]])-1, 
   main="Outdegree Histogram")

# Centrality
cat("Betweenness Centrality:", "\n")
print(betweenness(subgraphs[[4]]))
cat(mean(is.nan(betweenness(subgraphs[[4]]))), "of the values are NaN", "\n")

cat("Closeness Centrality:", "\n")
clr <- centr_clo(subgraphs[[4]], mode="all", normalized=T)
print(centr_clo(subgraphs[[4]], mode="all", normalized=T))
cat(mean(is.nan(closeness(subgraphs[[4]]))), "of the values are NaN", "\n")
sort(clr$res, decreasing = T, index.return = T)

cat("Local Transitivity:", "\n")
print(transitivity(subgraphs[[4]], type="local"))
```

### Interpretations

Density is the proportion of edges in a network compared to how many possible edges could exist. A density of `r edge_density(subgraphs[[4]])` indicates a sparse network with a low proportion of connected nodes. This is reasonable for a Google web graph, as only somewhat-related websites would have links connecting one to the other. 

Reciprocity is the proportion of edges from node A to B that have a corresponding edge traveling from B to A. This quantity tends to be higher in friendship networks where connections are mutual, and lower in follower networks due to their one-sided nature. Our value of `r reciprocity(subgraphs[[4]])` suggests a moderately low percentage of websites with mutual hyperlinks. 
The dyad census takes all pairs of nodes and calculates the number of mutual connections, one-sided (asymmetric) connections, and null connections among all node pairs. We see that there are `r dyad_census(subgraphs[[4]])$mut` mutually connected nodes and almost 20000 unconnected nodes. This echoes the low density found from earlier, and is perfectly reasonable since hyperlinks should be relevant to the website it was found on. 

Diameter measures the length of all shortest paths between nodes in the network and returns the maximum value, while average distance returns the mean. Higher values indicate more indirect connections. A mean distance of `r mean_distance(subgraphs[[4]])` is quite low, and suggests that connections seldom occur through a large number of intermediary websites.

Global transitivity is the likelihood that two different neighbors of a node also share an edge. High transitivity indicates a network with tight-knit groups and more clustering. This is the case for our network, as a transitivity of `r transitivity(subgraphs[[4]])` shows the close-knit nature of websites with hyperlinks conencting each other, even on a global scale. 

Degree is a node-level measure of how many edges connected to it. Since we are working with a directed graph, the number of in-edges and out-edges are distinct. There is a clear right skew in both histograms, which is expected for a web graph since most websites will have a handful of connections but few will have a great deal.

Betweenness is a centrality measure that finds the frequency of a node existing within the shortest path between two other nodes. Nodes with high betweenness centrality are often critical points in the network that control information flow. In our network some nodes have a betweenness measure of over 1400 while others have a betweenness of 0. Pages like Yahoo or trusted encyclopedia pages likely had these high centralities, since they would have been referred to frequently by other sites even if they were not directly related.  

Closeness is a centrality measure that measures the distance from a node to all other nodes. Important nodes will have a higher closeness since they interact with the other nodes more closely. The closeness values typically fall between 0.15 and 0.25, meaning that websites take around 4-7 hyperlinks to jump to the next website. The theme of close-knittedness between websites persists here; website hyperlinks will typically be relevant to the original website and most connections will not stray too far away. The low centralization of <0.15 suggests that centrality is distributed fairly evenly across the network, with very few websites dominating traffic.   

Local transitivity is the smaller-scale measure of transitivity, and tells whether the data's clustering occurs frequently between small groups. Values frequently take values close to 1, which again shows that web graphs are tightly connected on a local level. Sites covering related topics will likely reference each other. 




  
  

## Community Detection

Because the community detection algorithms want an undirected graph and our current graph is directed, we need to convert it into an undirected graph first. Then, we can run all of the community detection algorithms on each subgraph and store it all inside one dataframe. We will also store the results inside a comunity objects data structure so that we can use the results later.

```{r}
# convert all into undirected graphs
undirected_subgraphs <- vector("list", 4)
for (i in 1:4){
  undirected_subgraphs[[i]] <- as.undirected(subgraphs[[i]], mode="collapse")
}
# list to store all results
all_results <- list()
community_objects <- list()

for (i in 1:4) {
  g <- undirected_subgraphs[[i]]
  
  community_objects[[i]] <- list(
    fast_greedy = cluster_fast_greedy(g),
    louvain     = cluster_louvain(g),
    walktrap    = cluster_walktrap(g)
  )
  
  # Grab the objects we just saved to calculate stats
  objs <- community_objects[[i]]
  
  all_results[[i]] <- data.frame(
    Subgraph      = i,
    Algorithm     = c("Fast Greedy", "Louvain", "Walktrap"),
    N_communities = c(length(objs$fast_greedy), length(objs$louvain), length(objs$walktrap)),
    Modularity_Q  = round(c(modularity(objs$fast_greedy), modularity(objs$louvain), modularity(objs$walktrap)), 4),
    Max_Comm_Size = c(max(sizes(objs$fast_greedy)), max(sizes(objs$louvain)), max(sizes(objs$walktrap)))
  )
}

# combine list into one table
final_comparison <- do.call(rbind, all_results)
final_comparison
```

Looking at the results, we see that Walktrap generally detects more communities than the other 2 algorithms. All of the algorithms have high modularity greater than 0.7, suggesting that they found clear communities.


## Adjacency Matrix
```{r}
for(i in 1:4){
  adj_mat <- as_adjacency_matrix(subgraphs[[i]], sparse = FALSE)
  
  image(adj_mat,
        main = paste("Subgraph", i, "- Original Node Order"),
        col = c("white", "steelblue"),
        axes = FALSE)
}
# convert all into undirected graphs
undirected_subgraphs <- vector("list", 4)
for (i in 1:4){
  undirected_subgraphs[[i]] <- as.undirected(subgraphs[[i]])
}
# list to store all results
all_results <- list()

for (i in 1:4) {
  g <- undirected_subgraphs[[i]]
  
  # run algs
  fg   <- cluster_fast_greedy(g)
  louv <- cluster_louvain(g)
  walk <- cluster_walktrap(g)
  
  # store into list
  all_results[[i]] <- data.frame(
    Subgraph      = i,
    Algorithm     = c("Fast Greedy", "Louvain", "Walktrap"),
    N_communities = c(length(fg), length(louv), length(walk)),
    Modularity_Q  = round(c(modularity(fg), modularity(louv), modularity(walk)), 4),
    Max_Comm_Size = c(max(sizes(fg)), max(sizes(louv)), max(sizes(walk)))
  )
}

# combine list into one table
final_comparison <- do.call(rbind, all_results)
final_comparison
```



## Adjacency Matrix
```{r}
for(i in 1:4){
  adj_mat <- as_adjacency_matrix(subgraphs[[i]], sparse = FALSE)
  
  image(adj_mat,
        main = paste("Subgraph", i, "- Original Node Order"),
        col = c("white", "steelblue"),
        axes = FALSE)
}
```


```{r}
set.seed(42)
for(i in 1:4){
  adj_mat <- as_adjacency_matrix(subgraphs[[i]], sparse = FALSE)
  
  kc_louv <- cluster_louvain(undirected_subgraphs[[i]])
  node_order <- order(membership(kc_louv))
  adj_mat_reordered <- adj_mat[node_order, node_order]
  
  image(adj_mat_reordered,
        main = paste("Subgraph", i, "- Reordered by Community"),
        col = c("white", "steelblue"),
        axes = FALSE)
}
```

The adjacency matrix of each 200-node subgraph was examined to identify structural patterns in the Google web graph. Each entry in the matrix represents the presence of a directed hyperlink between two pages. In the original node ordering, edges appear scattered with no discernible pattern, reflecting the arbitrary order in which nodes were sampled.
To reveal underlying structure, nodes were reordered by Louvain community membership. After reordering, clear diagonal block patterns emerged, particularly in Subgraph 4 which contained 1,608 edges. Each block along the diagonal represents a group of web pages that densely hyperlink to one another, likely corresponding to topically related pages. The sparsity off the diagonal indicates that pages from different communities rarely link directly to each other. This visual confirmation aligns with the community detection results, supporting the presence of meaningful cluster structure within the Google web graph.

## Reordering Nodes for Community Detection

As we see above, all of the nodes are in one connected component. As such, this step is essentially the same as above. We will reorder the adjacency matrices for all of the remaining subgraphs as well.

```{r}
for (i in 1:4) {
  A <- as_adjacency_matrix(undirected_subgraphs[[i]], sparse = FALSE)
  
  par(mfrow = c(1, 3))
  
  # loop through the algorithms
  alg_names <- names(community_objects[[i]])
  
  for (alg in alg_names) {
    comm_obj <- community_objects[[i]][[alg]]
    
    ord <- order(membership(comm_obj))
    
    A_reordered <- A[ord, ord]
    
    image(A_reordered,
          main = paste("Sub", i, ":", alg),
          xlab = "Node (reordered)", 
          ylab = "Node (reordered)",
          col=c("white", "steelblue"),
          axes = FALSE)
  }
  
  # reset plots for next one
  par(mfrow = c(1, 1))
}
```

After reordering by community for each algorithm, we can see that the nodes form multiple boxes. This represents communities where nodes are highly interconnected. The adjacency matrix also happens to be mirrored across the y=x line, and that's because we used the undirected graph. We chose to use the undirected graph instead of the directed graph because that is the graph the algorithms used to detect the communities. If you also compare the adjencency matrix to the algorithm results, you see that the boxes correspond to how many communities each algorithm detected.

There are two main shapes we see in the adjacency matrix. First is this sparse L shape. This represents a central node that connects to many other nodes. In the context of Google, this could be a site like Wikipedia that has links to many different communities. Next is a solid box. This represents a community where all the nodes are closely interconnected. This could be a set of subpages, where all pages link to each other. 





# Results






# Conclusion



















\newpage

# References


\newpage

# Appendix

```{r appendix-code, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
