---
title: "SNA Final Project Slides"
author: "Richard Zhang, Toby Liu, Frank Wang, Max Peng"
subtitle: PSTAT 194CS
output:
  ioslides_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(igraph)
```

```{r load-data, include=FALSE}
library(igraph)
edges <- read.table("web-Google.txt.gz", comment.char = "#")
graph <- graph_from_data_frame(edges, directed = TRUE)

set.seed(123)
subgraphs <- list()
for(i in 1:4){
  start_node <- sample(V(graph), 1)
  walk <- random_walk(graph, start = start_node, steps = 500, mode = "all")
  nodes_sampled <- unique(walk)[1:200]
  subgraphs[[i]] <- induced_subgraph(graph, nodes_sampled)
}

undirected_subgraphs <- lapply(subgraphs, function(g) as.undirected(g, mode = "collapse"))
```

# Google Dataset

## Dataset Characteristics:
- Type: Directed web graph
- Nodes (webpages): 875713
- Edges (hyperlinks): 5105039
- Source: Google Programming Contest (2002)

## Step 6 - Adjacency Matrix

- 200x200 matrix where each cell = 1 if a hyperlink exists between two pages
- Original ordering: edges appear scattered, no visible structure
- Goal: reorder nodes to reveal hidden patterns

## Adjacency Matrix: Reordered by Community

- Nodes reordered by Louvain community membership
- Diagonal blocks emerge — pages in the same community densely hyperlink each other
- Subgraph 4 (1,608 edges) shows the clearest block structure
- Off-diagonal sparsity confirms pages from different communities rarely link directly

## Step 6 - Adjacency Matrix Reordered
```{r, fig.height=5, fig.width=5, fig.align='center'}
adj_mat <- as_adjacency_matrix(subgraphs[[4]], sparse = FALSE)
kc_louv <- cluster_louvain(undirected_subgraphs[[4]])
node_order <- order(membership(kc_louv))
adj_mat_reordered <- adj_mat[node_order, node_order]
image(adj_mat_reordered,
      col = c("white", "steelblue"),
      axes = FALSE)
```