---
title: "PSTAT194CS_SNA_Final_Project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,   # hides code
  message = FALSE,
  warning = FALSE
)
```

# Dataset

# Loading and Preparing Data
```{r}
url = "https://snap.stanford.edu/data/web-Google.txt.gz"
download.file(url, destfile = "web-Google.txt.gz", mode = "wb")
```
```{r}
library(igraph)

edges = read.table(
  "web-Google.txt.gz",
  comment.char = "#",
)

graph = graph_from_data_frame(edges, directed=TRUE)
```

# Understanding our Network
```{r}
cat("Number of Nodes:",vcount(graph),"\n")
cat("Number of Edges:",ecount(graph))
```

```{r}
is_directed(graph)
```



```{r}
is_weighted(graph)
```


```{r}
summary(graph)
```






# Creating Subgraphs of 200-nodes



## Code Implemenation
```{r}
set.seed(123)

sub.network = lapply(1:4, function(x){
  induced_subgraph(graph, sample(V(graph),200))
})
```

## Checking Subgraphs
```{r}
sub.network[[1]]
sub.network[[2]]
sub.network[[3]]
sub.network[[4]]
```










## Random Walk New Implementation
```{r}
set.seed(123)

subgraphs <- list()

for(i in 1:4){

  start_node <- sample(V(graph), 1)

  walk <- random_walk(graph,
                      start = start_node,
                      steps = 500,
                      mode = "all")

  nodes_sampled <- unique(walk)

  nodes_sampled <- nodes_sampled[1:200]

  subgraphs[[i]] <- induced_subgraph(graph, nodes_sampled)

}

```

## Checking our Subgraphs
```{r}
subgraphs[[1]]
subgraphs[[2]]
subgraphs[[3]]
subgraphs[[4]]
```


# Visualization 

```{r}
layouts = list(layout_with_fr, layout_with_kk, layout_in_circle)

for(i in 1:4){
  for (j in 1:3){
    plot(
      subgraphs[[i]],
      layout = layouts[[j]],
      vertex.size = 4,
      vertex.label = NA,
      edge.arrow.size = 0.2,
      main= paste("Subgrah", i,"- Layout", j)
    )
  }
}
```

# Metrics Evaluated for ALL 4 SUBGRAPHS (probably don't do this)

```{r, eval = TRUE}
for (i in 1:4) {
  cat("Metrics for Subgraph", i, ":\n", "NETWORK LEVEL METRICS", "\n")
  
  cat("Density:", "\n") 
  print(edge_density(subgraphs[[i]]))

  # Reciprocity
  cat("Reciprocity:", "\n")
  print(reciprocity(subgraphs[[i]]))

  # Dyad Census 
  cat("Pairwise Relationships:", "\n")
  print(dyad_census(subgraphs[[i]]))

  # Transitivity 
  cat("Transitivity:", "\n")
  print(transitivity(subgraphs[[i]]))
  
  # Average distance between nodes
  cat("Mean Distance Between Nodes:", "\n")
  print(mean_distance(subgraphs[[i]]))

  # Diameter
  cat("Diameter and the Associated Nodes:", "\n")
  print(diameter(subgraphs[[i]], weights = NA))
  print(farthest_vertices(subgraphs[[i]]))
  
  # NODE LEVEL METRICS
  cat("NODE LEVEL METRICS", "\n", "Degree:", "\n")
  print(degree(subgraphs[[i]], mode = "in"))
  
  # Centrality
  cat("Betweenness Centrality:", "\n")
  print(betweenness(subgraphs[[i]]))
  cat(mean(is.nan(betweenness(subgraphs[[i]]))), "of the values are NaN", "\n")
  
  cat("Closeness Centrality:", "\n")
  clr <- centr_clo(subgraphs[[i]], mode="all", normalized=T)
  print(centr_clo(subgraphs[[i]], mode="all", normalized=T))
  cat(mean(is.nan(closeness(subgraphs[[i]]))), "of the values are NaN", "\n")
  sort(clr$res, decreasing = T, index.return = T) # REMOVE ME Can remove if we don't want to plot
  
  cat("Local Transitivity:", "\n")
  print(transitivity(subgraphs[[i]], type="local"))

  # Edge Level
  edge_betweenness(subgraphs[[i]])
}
```



# 5) Community Detection

Because the graph is directed, we need to convert it into an undirected graph for community detection. 
```{r}
# convert all into undirected graphs
undirected_subgraphs <- vector("list", 4)
for (i in 1:4){
  undirected_subgraphs[[i]] <- as.undirected(subgraphs[[i]])
}

for (i in 1:4){
  kc_fg <- cluster_fast_greedy(undirected_subgraphs[[i]])
  cat("Algorithm: Fast Greedy\n")
  cat("Number of communities:", length(kc_fg), "\n")
  cat("Modularity Q =", round(modularity(kc_fg), 4), "\n")
  cat("Community sizes:\n")
  print(sizes(kc_fg))
  
  kc_louv <- cluster_louvain(undirected_subgraphs[[i]])
  cat("Algorithm: Louvain\n")
  cat("Number of communities:", length(kc_louv), "\n")
  cat("Modularity Q =", round(modularity(kc_louv), 4), "\n")
  cat("Community sizes:\n")
  print(sizes(kc_louv))
  
  kc_walk <- cluster_walktrap(undirected_subgraphs[[i]])
  cat("Algorithm: Walktrap\n")
  cat("Number of communities:", length(kc_walk), "\n")
  cat("Modularity Q =", round(modularity(kc_walk), 4), "\n")
  cat("Community sizes:\n")
  print(sizes(kc_walk))
}

```































# 5) Community Detection

Because the graph is directed, we need to convert it into an undirected graph for community detection. 
```{r}
# convert all into undirected graphs
undirected_subgraphs <- vector("list", 4)
for (i in 1:4){
  undirected_subgraphs[[i]] <- as.undirected(subgraphs[[i]])
}

for (i in 1:4){
  kc_fg <- cluster_fast_greedy(undirected_subgraphs[[i]])
  cat("Algorithm: Fast Greedy\n")
  cat("Number of communities:", length(kc_fg), "\n")
  cat("Modularity Q =", round(modularity(kc_fg), 4), "\n")
  cat("Community sizes:\n")
  print(sizes(kc_fg))
  
  kc_louv <- cluster_louvain(undirected_subgraphs[[i]])
  cat("Algorithm: Louvain\n")
  cat("Number of communities:", length(kc_louv), "\n")
  cat("Modularity Q =", round(modularity(kc_louv), 4), "\n")
  cat("Community sizes:\n")
  print(sizes(kc_louv))
  
  kc_walk <- cluster_walktrap(undirected_subgraphs[[i]])
  cat("Algorithm: Walktrap\n")
  cat("Number of communities:", length(kc_walk), "\n")
  cat("Modularity Q =", round(modularity(kc_walk), 4), "\n")
  cat("Community sizes:\n")
  print(sizes(kc_walk))
}

```












































\newpage

## Appendix

```{r appendix-code, ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
